Paper 1 Notebook

### 2024-07-05 minos/outcomes/paper_1_notebook.Rmd
updates to notebook for paper revision

```{r "preamble"}
source("minos/transitions/utils.R")
source("minos/transitions/transition_model_functions.R")
source("minos/transitions/paper_1_plots.R")
library(dplyr)
library(texreg)
library(lme4)
library(glmnet)
library(mice)
library(sjPlot)
library(visreg)
library(ggplot2)
library(ggpattern)
library(VIM)
library(caret)
library(tidyverse)
library(performance)
library(broom.mixed)
library(ggridges)
library(ggmice)
library(MASS)
save.path <- "minos/outcomes/paper1_plots/"
raw.data.dir <- "data/raw_US/"
mice.data.dir <- "data/composite_US/"
final.data.dir <- mice.data.dir
```

@@

# get raw data file names from 2011-2013.
filelist <- list.files(raw.data.dir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[21:23] # get files for 2011-2013.
# load raw data.
raw.data <- do.call(rbind, lapply(filelist, read.csv))

# baseline OLS model formula. 
raw_model_formula_string <- "SF_12 ~ factor(sex) + 
                  relevel(factor(ethnicity), ref='WBI') + 
                  scale(age) + 
                  factor(education_state) + 
                  factor(labour_state) + 
                  relevel(factor(NSSEC), ref=1) +
                  relevel(factor(region), ref='London') +
                  scale(hh_income)"

@@

# get only needed variables and replace missing codes with NA.
raw.data <- update_column_names(raw.data)
raw.data <- raw.data[, append(variables, c("time", 'pidp', 'weight'))]
raw.data <- replace.missing(raw.data)

@@

# get raw data file names.
full.raw.filelist <- list.files(raw.data.dir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[21:30] # get files for 2011-2013.
# load raw data.
full.raw.data <- do.call(rbind, lapply(full.raw.filelist, read.csv))
full.raw.data <- update_column_names(full.raw.data)

# get only needed variables and replace missing codes with NA.
full.raw.data <- full.raw.data[, append(variables, c("time", 'pidp', 'weight'))]
full.raw.data <- replace.missing(full.raw.data)
```

@@

# create save.path if it doesn't exist and save OLS coefficients as a latex table there. 
create.if.not.exists(save.path)
write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF12_OLS_coefficients.txt')
  tex_label <- paste0("table: baseline_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("Baseline SF_12 OLS"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "ols"))    
  texreg(sf12.lm, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='small') # small font
}

# diagnostic plots for OLS presentation. 
forest_plot_lm(sf12.lm, paste0(save.path, "presentation_forest.pdf"))
qq_plot(resid(sf12.lm), paste0(save.path, "presentation_qq.pdf"))
residual_density_plot(res=resid(sf12.lm), file_name=paste0(save.path, "presentation_residual_density.pdf"), guide="normal")

@@

start.data <- raw.data[which(raw.data$time==2013),]
start.data.colnames <- colnames(start.data)
start.data.colnames[which( colnames(start.data)=="job_sec" )] <- "NSSEC"
start.data.colnames[which( colnames(start.data)=="hh_netinc" )] <- "hh_income"
start.data.colnames[which( colnames(start.data)=="labour_state_raw" )] <- "labour_state"
colnames(start.data) <- start.data.colnames
handover.data <- start.data[, c("pidp", "SF_12", "time")]
rolling.data <- start.data

for (i in 1:7){
  start.data$SF_12 <- predict(sf12.lm, rolling.data)
  start.data[, 'time'] <- start.data[, 'time'] +1
  handover.data <- rbind(handover.data, start.data[, c("pidp", "SF_12", "time")])
}

@@

@@

```{r}
# load in all 2009-2020 datasets
mice.filelist <- list.files(mice.data.dir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')
mice.data <- do.call(rbind, lapply(mice.filelist, read.csv))

@@

# missingness MICE model formula
mice_formula_string <- "SF_12 ~ 
                        factor(sex) + 
                        relevel(factor(ethnicity), ref='WBI') + 
                        scale(age) + 
                        factor(education_state) + 
                        factor(S7_labour_state) + 
                        relevel(factor(NSSEC), ref=1) +
                        relevel(factor(region), ref='London') +
                        scale(hh_income) + 
                        scale(I(hh_income**2)) +
                        factor(housing_quality) +
                        factor(neighbourhood_safety) + 
                        factor(loneliness) +
                        ncigs +
                        nutrition_quality"

@@

# preserving key columns that aren't imputed for later. cant be included. in imputation set due to perfect seperation. 
mice.pidp <- mice.data[which(mice.data$time %in% c(2011, 2012, 2013)), "pidp"]
mice.time <- mice.data[which(mice.data$time %in% c(2011, 2012, 2013)), "time"]
mice.weights <- mice.data[which(mice.data$time %in% c(2011, 2012, 2013)), "weight"]

mice.2018.pidp <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "pidp"]
mice.2018.time <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "time"]
mice.2018.weights <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "weight"]

lm.no.mice <- lm(mice_formula_string, data=mice.data)
forest_plot(lm.no.mice,  paste0(save.path, "no_mice_forest.pdf"))
```

@@

method <- c("pmm",
            'pmm',
            'pmm',
            'pmm',#'polr',
            'pmm',
            'pmm', #'polr',
            'pmm',
            'pmm', #'polr', #'pmm',
            'pmm',
            'pmm',
            'pmm',
            'pmm',
            'pmm',
            'pmm',
            "pmm",
            "pmm")

@@

mice.data$ethnicity <- factor(mice.data$ethnicity)
mice.data$S7_labour_state <- factor(mice.data$S7_labour_state)
mice.data$region <- factor(mice.data$region)
mice.data$loneliness <- factor(mice.data$loneliness)
#mice.data$heating <- factor(mice.data$heating)
#mice.data$job_sec <- factor(mice.data$job_sec)
#mice.data$education_state <- factor(mice.data$education_state)
#mice.data$sex <- factor(mice.data$sex)
#mice.data$smoker <- factor(mice.data$smoker)
#mice.data$behind_on_bills <- factor(mice.data$behind_on_bills)
#mice.data$ghq_depression <- factor(mice.data$ghq_depression)
#mice.data$ghq_happiness <- factor(mice.data$ghq_happiness)
#mice.data$clinical_depression <- factor(mice.data$clinical_depression)
#mice.data$scsf1 <- factor(mice.data$scsf1)
#mice.data$health_limits_social <- factor(mice.data$ghq_depression)
#mice.data$housing_tenure <- factor(mice.data$housing_tenure)
#mice.data$urban <- factor(mice.data$urban)
#mice.data$marital_status <- factor(mice.data$marital_status)

load_cache <- T
if (load_cache) {
  mice_set <- readRDS("data/transitions/MICE_set.rds")
} else {
  n_iter <- 30
  max_iter <- 10
  ignore_cols = c("birth_year")
  # future mice is parallelised version of MICE.
  mice_set <- mice(data = mice.data[, imp_columns],
                   m = n_iter, maxit = max_iter,
                   remove.collinear=T)

@@

pre.mice.lm <- lm(mice_formula_string, data=mice.data, weights = weight)
forest_plot(pre.mice.lm , paste0(save.path,'mice_SF12_forest.pdf'))
qq_plot(resid(pre.mice.lm ), paste0(save.path, "mice_qq.pdf"))

# writing coefficients table for MICE. 
# is there a version for pool table?
write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF12_without_MICE_OLS_coefficients.txt')
  tex_label <- paste0("table: without_MICE_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("Without MICE SF_12 OLS"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "without_mice"))    
  texreg(mice.lm.object, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='small') # small font

  # same again for MICE pool. 
  texreg_file <- paste0(save.path, 'SF12_MICE_pool_OLS_coefficients.txt')
  tex_label <- paste0("table: MICE_pool_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("MICE SF_12 OLS Pool"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "ols"))    
  texreg(final.pool, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='small') # small font
}

@@

# only looking at 2011-2013 subset of data. convert to data frame, subset, and back to MICE mids. 
# also adding variables that can't be imputed back in. pidp time weight.
mice.long <- complete(mice_set, action = "long", include=T)
#mice.long.subset <- mice.long[which(mice.long$time %in% c(2011, 2012, 2013)), ]

mice.long$pidp <- mice.data$pidp
mice.long$weight <- mice.data$weight
mice.long$time <- mice.data$time
mice.long <- update_column_names(mice.long)
mice_set.full <- as.mids(mice.long)

mice.long.subset <- mice.long[which(mice.long$time %in% c(2011, 2012, 2013)), ]
mice_set.subset <- as.mids(mice.long.subset)

@@

# coefficient forest plot. 
forest_plot(mice.lm.object, paste0(save.path,'mice_SF12_forest.pdf'))
qq_plot(resid(mice.lm.object), paste0(save.path, "mice_qq.pdf"))

# writing coefficients table for MICE. 
# is there a version for pool table?
write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF12_MICE_OLS_coefficients.txt')
  tex_label <- paste0("table: MICE_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("MICE SF_12 OLS"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "ols"))    
  texreg(mice.lm.object, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='small') # small font

  # same again for MICE pool. 
  texreg_file <- paste0(save.path, 'SF12_MICE_pool_OLS_coefficients.txt')
  tex_label <- paste0("table: MICE_pool_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("MICE SF_12 OLS Pool"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "ols"))    
  texreg(final.pool, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='small') # small font
}

#screenreg final.pool
```

@@

```{r cross validated nowcasting}
# split data into five
# transition models
# compare coefficients in plots.

@@

@@

write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF12_cross_val_OLS_coefficients.txt')
  tex_label <- paste0("table: ols_cross_val_coefficients") 
  tex_headers <- c(texreg::names2latex("5-Fold Cross Validated SF_12 OLS"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("cv_SF_12", "ols"))    
  texreg(summary(cross.val.sf12.model), # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='small') # small font
}
```

@@

```{r mice_handovers}
start.data <- crossval.mice.data[which(crossval.mice.data$time==2013),]
rolling.data <- start.data#[, c("pidp", "SF_12", "time")]
handover.data <- rolling.data[, c("pidp", "SF_12", "time")]
for (i in 1:7){
  rolling.data$SF_12 <- predict(pooled_lm, newdata = rolling.data)
  start.data$SF_12 <- rolling.data$SF_12
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data[, c("pidp", "SF_12", "time")])
}

@@

```{r cross_val_mice_handovers}
start.data <- crossval.mice.data[which(crossval.mice.data$time==2013),]
rolling.data <- start.data
handover.data <- rolling.data[, c("pidp", "SF_12", "time")]

for (i in 1:7){
  rolling.data$SF_12 <- predict(cross.val.sf12.model, rolling.data)
  start.data$SF_12 <- rolling.data$SF_12
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data[, c("pidp", "SF_12", "time")])
}

@@

texreg(c(lasso.min, lasso.1se), # what model to save coefs for. 
       file=texreg_file,  # where to save.
       custom.model.names = tex_headers, # title of table.
       caption=tex_caption, # \caption command option. 
       label = tex_label, # \label command option. 
       dcolumn=T,# nice column alignment. 
       booktabs=T, # nice hlines. (recommended)
       tabular=T,# tabulr env.
       single.row = T, # nicer one row formatting.
       longtable=T, # allows tables to wrap over pages
       use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
       fontsize='small') # small font

@@

>>>>

```{r setup}
knitr::opts_knit$set(root.dir = "/home/rob/Minos")
getwd()
```

```{r "preamble"}
#setwd("../..")
library(here)
source(here::here("minos", "transitions", "utils.R"))
source(here::here("minos", "transitions", "transition_model_functions.R"))
source(here::here("minos", "transitions", "paper_1_plots.R"))
library(dplyr)
library(texreg)
library(lme4)
library(glmnet)
library(mice)
library(sjPlot)
library(visreg)
library(ggplot2)
library(ggpattern)
library(VIM)
library(caret)
library(tidyverse)
library(performance)
library(broom.mixed)
library(ggridges)
library(ggmice)
library(MASS)
library(qpcR)
library(cv)
library(Rcpp)

save.path <- "minos/outcomes/paper1_plots/"
#raw.data.dir <- "data/raw_US/"
raw.data.dir  <- "data/composite_US/"
mice.data.dir <- "data/composite_US/"
final.data.dir <- mice.data.dir
```

@@

#TODO RMSE AND PRESS STATS FOR CROSS VAL.  RMSE(predictions, test_data$outcomeName)
#TODO MAKE SURE CROSS VAL ON BOTH STAGES.
# get raw data file names from 2011-2013.
filelist <- list.files(raw.data.dir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[3:5]#[21:23] # get files for 2011-2013.
# load raw data.
raw.data <- do.call(rbind, lapply(filelist, read.csv))

# baseline OLS model formula. 
raw_model_formula_string <- "I(log((80 -SF_12)+10)) ~ factor(sex) + 
                  relevel(factor(ethnicity), ref='WBI') + 
                  scale(age) + 
                  factor(education_state) + 
                  factor(labour_state) + 
                  relevel(factor(NSSEC), ref=1) +
                  relevel(factor(region), ref='London') +
                  scale(hh_income)"

@@

# get only needed variables and replace missing codes with NA.
raw.data <- update_column_names(raw.data)
raw.data <- raw.data[, append(variables, c("time", 'pidp','hidp', 'weight'))]
raw.data <- replace.missing(raw.data)

@@

# get raw data file names.
full.raw.filelist <- list.files(raw.data.dir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')[3:13] #[21:30] # get files for 2011-2013.
# load raw data.
full.raw.data <- do.call(rbind, lapply(full.raw.filelist, read.csv))
full.raw.data <- update_column_names(full.raw.data)

# get only needed variables and replace missing codes with NA.
full.raw.data <- full.raw.data[, append(variables, c("time", 'pidp', 'hidp', 'weight'))]
full.raw.data <- replace.missing(full.raw.data)
```

@@

# create save.path if it doesn't exist and save OLS coefficients as a latex table there. 
create.if.not.exists(save.path)
write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF12_OLS_coefficients.txt')
  tex_label <- paste0("table: baseline_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("Baseline SF_12 OLS"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "ols"))    
  texreg(sf12.lm, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='tiny') # tiny font
}

# diagnostic plots for OLS presentation. 
forest_plot_lm(sf12.lm, paste0(save.path, "presentation_forest.pdf"))
log_forest_plot(sf12.lm, paste0(save.path, "presentation_forest.pdf"))
qq_plot(resid(sf12.lm), paste0(save.path, "presentation_qq.pdf"))
residual_density_plot(res=resid(sf12.lm), file_name=paste0(save.path, "presentation_residual_density.pdf"), guide="normal")

@@

start.data <- raw.data[which(raw.data$time==2013),]
start.data.colnames <- colnames(start.data)
start.data.colnames[which( colnames(start.data)=="job_sec" )] <- "NSSEC"
start.data.colnames[which( colnames(start.data)=="hh_netinc" )] <- "hh_income"
start.data.colnames[which( colnames(start.data)=="labour_state_raw" )] <- "labour_state"
colnames(start.data) <- start.data.colnames
handover.data <- start.data[, c("pidp", "SF_12", "hidp", "time")]
rolling.data <- start.data

for (i in 1:8){
  rolling.data$SF_12 <- (80-(exp(predict(sf12.lm, rolling.data))-10))
#  start.data$SF_12 <- predict(sf12.lm, rolling.data)
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data[, c("pidp", "hidp", "SF_12", "time")])
}

@@

```{r raw cross val}
# Baseline cross val.

raw.cross.val.sf12.model <- train(as.formula(raw_model_formula_string), data = raw.data,
               method = "lm",
               trControl = trControl)

print(RMSE(raw.cross.val.sf12.model))

#print(PRESS(raw.cross.val.sf12.model))
```

@@

```{r mice preprocessing}
# load in all 2009-2020 datasets
mice.filelist <- list.files(mice.data.dir,
                       include.dirs = FALSE,
                       full.names = TRUE,
                       pattern = '[0-9]{4}_US_cohort.csv')
mice.data <- do.call(rbind, lapply(mice.filelist, read.csv))

@@

# missingness MICE model formula
mice_formula_string <- "I(log(SF_12+10)) ~ 
                        factor(sex) + 
                        relevel(factor(ethnicity), ref='WBI') + 
                        scale(age) + 
                        factor(education_state) + 
                        factor(labour_state) + 
                        relevel(factor(NSSEC), ref=1) +
                        relevel(factor(region), ref='London') +
                        scale(hh_income) + 
                        scale(I(hh_income**2)) +
                        factor(housing_quality) +
                        factor(neighbourhood_safety) + 
                        factor(loneliness) +
                        ncigs +
                        nutrition_quality"

@@

# preserving key columns that aren't imputed for later. cant be included. in imputation set due to perfect seperation. 
mice.pidp <- mice.data[which(mice.data$time %in% c(2011, 2012, 2013)), "pidp"]
mice.hidp <- mice.data[which(mice.data$time %in% c(2011, 2012, 2013)), "hidp"]
mice.time <- mice.data[which(mice.data$time %in% c(2011, 2012, 2013)), "time"]
mice.weights <- mice.data[which(mice.data$time %in% c(2011, 2012, 2013)), "weight"]

mice.2018.pidp <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "pidp"]
mice.2018.hidp <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "hidp"]
mice.2018.time <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "time"]
mice.2018.weights <- mice.data[which(mice.data$time %in% c(2016, 2017, 2018)), "weight"]

lm.no.mice <- lm(mice_formula_string, data=mice.data)
forest_plot(lm.no.mice,  paste0(save.path, "no_mice_forest.pdf"))
log_forest_plot(lm.no.mice,  paste0(save.path, "no_mice_forest.pdf"))
```

@@

@@

mice.data$ethnicity <- factor(mice.data$ethnicity)
mice.data$labour_state <- factor(mice.data$labour_state)
mice.data$region <- factor(mice.data$region)
mice.data$loneliness <- factor(mice.data$loneliness)
#mice.data$heating <- factor(mice.data$heating)
#mice.data$job_sec <- factor(mice.data$job_sec)
#mice.data$education_state <- factor(mice.data$education_state)
#mice.data$sex <- factor(mice.data$sex)
#mice.data$smoker <- factor(mice.data$smoker)
#mice.data$behind_on_bills <- factor(mice.data$behind_on_bills)
#mice.data$ghq_depression <- factor(mice.data$ghq_depression)
#mice.data$ghq_happiness <- factor(mice.data$ghq_happiness)
#mice.data$clinical_depression <- factor(mice.data$clinical_depression)
#mice.data$scsf1 <- factor(mice.data$scsf1)
#mice.data$health_limits_social <- factor(mice.data$ghq_depression)
#mice.data$housing_tenure <- factor(mice.data$housing_tenure)
#mice.data$urban <- factor(mice.data$urban)
#mice.data$marital_status <- factor(mice.data$marital_status)

load_cache <- T
if (load_cache) {
  mice_set <- readRDS("data/transitions/MICE_set.rds")
} else {
  #n_iter <- 30
  #max_iter <- 10
  n_iter <- 1
  max_iter <- 1
  ignore_cols = c("birth_year")
  # future mice is parallelised version of MICE.
  mice_set <- mice(data = mice.data[mice.data$time %in%  c(2013, 2020), imp_columns],
                   m = n_iter, maxit = max_iter,
                   remove.collinear=T)

@@

pre.mice.lm <- lm(mice_formula_string, data=mice.data, weights = weight)
forest_plot(pre.mice.lm , paste0(save.path,'mice_SF12_forest.pdf'))
log_forest_plot(pre.mice.lm , paste0(save.path,'mice_SF12_forest.pdf'))
qq_plot(resid(pre.mice.lm ), paste0(save.path, "mice_qq.pdf"))
```

@@

# only looking at 2011-2013 subset of data. convert to data frame, subset, and back to MICE mids. 
# also adding variables that can't be imputed back in. pidp time weight.
mice.long <- complete(mice_set, action = "long", include=T) #
#mice.long$SF_12 <- mice.long$SF_12 + 10
#mice.long.subset <- mice.long[which(mice.long$time %in% c(2011, 2012, 2013)), ]

mice.long$pidp <- mice.data$pidp
mice.long$hidp <- mice.data$hidp
mice.long$weight <- mice.data$weight
mice.long$time <- mice.data$time
mice.long <- update_column_names(mice.long)
mice_set.full <- as.mids(mice.long)

mice.long.subset <- mice.long[which(mice.long$time %in% c(2011, 2012, 2013)), ]

as.mids2 <- function (data, .imp = 1, .id = 2) 
    {
      # paste added to get numerical value if .imp is factor
      ini <- mice(data[data[, .imp] == 0, -c(.imp, .id)], m = max(as.numeric(paste(data[, .imp]))), maxit = 0)
      names <- names(ini$imp)
          if (!is.null(.id)) {
            rownames(ini$data) <- data[data[, .imp] == 0, .id]
          }
          for (i in 1:length(names)) {
            # -1 removed and paste added to get numerical value if .imp is factor
            for (m in 1:(max(as.numeric(paste(data[, .imp]))))) {
              if (!is.null(ini$imp[[i]])) {
                indic <- data[, .imp] == m & is.na(data[data[, .imp] == 0, names[i]])
                ini$imp[[names[i]]][m] <- data[indic, names[i]]
          }
        }
      }
      return(ini)
    }
mice_set.subset <- as.mids2(mice.long.subset) # bug in as.mids can't handle this. nice fix here. https://stats.stackexchange.com/questions/138769/something-wrong-with-as-mids-from-mice-package-in-r

@@

# coefficient forest plot. 
forest_plot(mice.lm.object, paste0(save.path,'mice_SF12_forest.pdf'))
log_forest_plot(mice.lm.object, paste0(save.path,'mice_SF12_forest.pdf'))
qq_plot(resid(mice.lm.object), paste0(save.path, "mice_qq.pdf"))

# writing coefficients table for MICE. 
# is there a version for pool table?
write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF12_MICE_OLS_coefficients.txt')
  tex_label <- paste0("table: MICE_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("MICE SF_12 OLS"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "ols"))    
  texreg(mice.lm.object, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='tiny') # tiny font

  # same again for MICE pool. 
  texreg_file <- paste0(save.path, 'SF12_MICE_pool_OLS_coefficients.txt')
  tex_label <- paste0("table: MICE_pool_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("MICE SF_12 OLS Pool"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "ols"))    
  texreg(final.pool, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='tiny') # tiny font
}

# writing coefficients table for MICE. 
# is there a version for pool table?
write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF12_without_MICE_OLS_coefficients.txt')
  tex_label <- paste0("table: without_MICE_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("Without MICE SF_12 OLS"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "without_mice"))    
  texreg(mice.lm.object, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='tiny') # tiny font

  # same again for MICE pool. 
  texreg_file <- paste0(save.path, 'SF12_MICE_pool_OLS_coefficients.txt')
  tex_label <- paste0("table: MICE_pool_OLS_coefficients") 
  tex_headers <- c(texreg::names2latex("MICE SF_12 OLS Pool"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("SF_12", "ols"))    
  texreg(final.pool, # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='tiny') # tiny font
}
#screenreg final.pool
```

@@

```{r mice cross validated nowcasting}
# split data into five
# transition models
# compare coefficients in plots.

@@

print(RMSE(cross.val.sf12.model))

@@

write_coefs <- T
if (write_coefs)
{
  texreg_file <- paste0(save.path, 'SF12_cross_val_OLS_coefficients.txt')
  tex_label <- paste0("table: ols_cross_val_coefficients") 
  tex_headers <- c(texreg::names2latex("5-Fold Cross Validated SF_12 OLS"))
  tex_caption <- texreg::names2latex(get_tex_table_caption("cv_SF_12", "ols"))    
  texreg(summary(cross.val.sf12.model), # what model to save coefs for. 
         file=texreg_file,  # where to save.
         custom.model.names = tex_headers, # title of table.
         stars = c(0.001, 0.01, 0.05), # p-value significance symbols. 
         digits=4, # p values significant figures.
         caption=tex_caption, # \caption command option. 
         label = tex_label, # \label command option. 
         dcolumn=T,# nice column alignment. 
         booktabs=T, # nice hlines. (recommended)
         tabular=T,# tabular env.
         single.row = T, # nicer one row formatting.
         longtable=T, # allows tables to wrap over pages
         use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
         fontsize='tiny') # tiny font
}
```

@@

```{r mice_handovers}
start.data <- crossval.mice.data[which(crossval.mice.data$time==2013),]
rolling.data <- start.data#[, c("pidp", "SF_12", "time")]
handover.data <- rolling.data[, c("pidp", "hidp", "SF_12", "time")]
for (i in 1:8){
  rolling.data$SF_12 <- exp(predict(pooled_lm, newdata = rolling.data))-10
  start.data$SF_12 <- rolling.data$SF_12
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data[, c("pidp", "hidp", "SF_12", "time")])
}

@@

```{r cross_val_mice_handovers}
start.data <- crossval.mice.data[which(crossval.mice.data$time==2013),]
rolling.data <- start.data
handover.data <- rolling.data[, c("pidp", "hidp", "SF_12", "time")]

for (i in 1:8){
  rolling.data$SF_12 <- exp(predict(cross.val.sf12.model, rolling.data))-10
  start.data$SF_12 <- rolling.data$SF_12
  rolling.data[, 'time'] <- rolling.data[, 'time'] +1
  handover.data <- rbind(handover.data, rolling.data[, c("pidp", 'hidp', "SF_12", "time")])
}

@@

texreg(c(lasso.min, lasso.1se), # what model to save coefs for. 
       file=texreg_file,  # where to save.
       custom.model.names = tex_headers, # title of table.
       caption=tex_caption, # \caption command option. 
       label = tex_label, # \label command option. 
       dcolumn=T,# nice column alignment. 
       booktabs=T, # nice hlines. (recommended)
       tabular=T,# tabulr env.
       single.row = T, # nicer one row formatting.
       longtable=T, # allows tables to wrap over pages
       use.packages=F, # don't add \usepackage commands to allow for direct importing of files as inputs
       fontsize='tiny') # tiny font

@@

```{r glmm_crossval}
